TODO: A Survey of StarCraft AI Techniques

### 2013-02-14 survey.tex
Fix some typos.

Figure \ref{fig:starcraft} shows a screenshot of Starcraft showing a player playing Terrans. In order to win a Starcraft game, players must first gather resources (minerals and gas), which they can spend in creating additional buildings and units. As resources become available, players need to allocate them for creating more buildings (which reinforce the economy, and allow players to create units or unlock stronger units) and train attack units. Units must be distributed to accomplish different tasks such as: reconnaissance, defense and attack. While performing all of those tasks, players also need to strategically understand the geometry of the map at hand, in order to decide where to place new buildings (concentrate in a single area, or expand to different areas) or where to set defensive outposts. Finally, when offensive units of two players meet, each player must quickly maneuver each of the units in order to fight the a battle, which requires quick and reactive control of each of the units in order to successfully fight a battle.

@@

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMS are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, strategic decision making is a hard problem, and these hard-coded approaches struggle to encode dynamic, adaptive behaviors.

@@

\begin{itemize}
\item BroodwarBotQ: uses abstraction for micro-management, and divide-and-conquer for macro-management and intelligence gathering. To avoid conflicts between modules (since the individual tasks of each of the modules are not completely independent), BBQ uses an arbitrator.
\item Nova: is similar in design as BroodwarBotQ, and uses abstraction for micro-management, and divide-and-conquer for macro-management. The differences are that Nova does not have an arbitrator to resolve conflicts, but has a higher-level module ({\em strategy manager}), which posts information to the blackboard that the rest of modules follow (thus, making use of abstraction).
\item UAlbertaBot: also uses abstraction in micro-management like the previous two bots. But it also uses it in macro-management: as can be seen, the production manager sends commands to the building manager, who is in charge of producing the buildings. This bot also uses divide-and-conquer, and tasks like scouting and resource gathering are managed by separate, independent modules.
\item Skynet: makes extensive use of both abstraction and divide-and-conquer. Considering the decision making component of Skynet, we can see a high level module that issues commands to a series of tactics modules. The collection of tactic modules queue {\em tasks} (that are analogous to the abstract actions used in SPAR). Each different task has a specific low level module that knows how to execute it. Thus, Skynet uses a 3 layered abstraction hierarchy, and uses divide-and-conquer in all levels except the highest.
\item SPAR: only uses abstraction. Its high-level module determines the strategy to use, and the tactical decision module divides it into a collection of {\em abstract actions}, that are executed by the lower-level modules.
\item AIUR: is mainly divide-and-conquer oriented, for macro as well as for micro, with a slight abstraction on macro with a SpendManager deciding how to spend resources among Base, Production and Construction Managers. At the beginning of a game, the MoodManager initializes a ``mood'' which will influence both micro and macro. For example, micro-management is divided into three independent managers: the {\em Defense Manager}, controlling military units when there is nothing special, the {\em Under Attack Manager}, activated when the opponent is attacking our bases, and the {\em Army Manager}, taking control of units when it is time to attack, following a timing given by the current mood.
\item BTHAI: uses a two-tier abstraction hierarchy, where a collection of high-level modules command a collection of lower-level agents in charge of each of the units. At the high-level, BTHAI uses divide-and-conquer, having multiple high-level modules issuing commands to the lower-level units.
\end{itemize}

@@

One interesting aspect of the five bots described above is that, while all of them are reactive at the lower level (reactive control), most if not all of them, are scripted at the highest level of abstraction. BTHAI reads build and squad formations from a predefined script, Nova's {\em Strategy Manager} is a predefined finite-state machine, BBQ's construction manager reads the build order from a predefined script, and Skynet's {\em BuildOrder Manager} is basically a predefined script. Such scripts describe the strategy that the bots will use, however, such strategy is always fixed. One could see this pre-scripting as if each bot defined a ``high-level programming language'' to describe Starcraft strategies, and the bots themselves are just interpreters of such strategy. Compared to current approaches for Chess or Go, this scripting seems a rigid and inflexible, but responds to the much higher complexity of the Starcraft game. An interesting exception to that is UAlbertaBot, which uses a search algorithm in the {\em Production Manager} to find near-optimal build orders. Another interesting case is AIUR, that uses a {\em Mood Manager} to randomly pick a mood among five (rush, aggressive, defensive, macro, fast expand), which will influence the build order, strategy and tactics, all of them scripted so far. This mood can change during the game, following the opponent behavior.

@@

In the end, Skynet\footnote{\url{http://code.google.com/p/skynetbot/}} won the competition with its solid Protoss play (results are summarized in Table \ref{tab:aiide2011}). The bot executed one of a small set of
strategies randomly at the start of the match based on the map and the race of the opponent. Skynet would
then amass a medium to large sized army and expand before moving out to attack. Good use of of Dragoon range
and kiting micromanagement allowed it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

As the CIG competition games were executed manually due to
a lack of available software (the AIIDE program was not yet
available at that time), we separated the ten entries into
two brackets. In each bracket of 5 bots, a round-robin
tournament was held with 10 repetitions per pairing, resulting
in 40 games per bot.
The 5 maps chosen for the first round were selected from the pool
of well-known league play maps found on the internet:
\emph{(2)MatchPoint 1.3}, \emph{(4)Fighting Spirit 1.3}, 
\emph{iCCupdestination 1.1}, \emph{iCCup gaia}, and 
\emph{iCCup great barrier reef}. Each bot pairing played
on every map twice, with switched starting positions.

>>>>

Figure \ref{fig:starcraft} shows a screenshot of Starcraft showing a player playing Terrans. In order to win a Starcraft game, players must first gather resources (minerals and gas), which they can spend in creating additional buildings and units. As resources become available, players need to allocate them for creating more buildings (which reinforce the economy, and allow players to create units or unlock stronger units) and train attack units. Units must be distributed to accomplish different tasks such as: reconnaissance, defense and attack. While performing all of those tasks, players also need to strategically understand the geometry of the map at hand, in order to decide where to place new buildings (concentrate in a single area, or expand to different areas) or where to set defensive outposts. Finally, when offensive units of two players meet, each player must quickly maneuver each of the units in order to fight a battle, which requires quick and reactive control of each of the units in order to successfully fight the battle.

@@

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMs are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, strategic decision making is a hard problem, and these hard-coded approaches struggle to encode dynamic, adaptive behaviors.

@@

\begin{itemize}
\item BroodwarBotQ: uses abstraction for micro-management, and divide-and-conquer for macro-management and intelligence gathering. To avoid conflicts between modules (since the individual tasks of each of the modules are not completely independent), BBQ uses an arbitrator.
\item Nova: is similar in design as BroodwarBotQ, and uses abstraction for micro-management, and divide-and-conquer for macro-management. The differences are that Nova does not have an arbitrator to resolve conflicts, but has a higher-level module ({\em strategy manager}), which posts information to the blackboard that the rest of modules follow (thus, making use of abstraction).
\item UAlbertaBot: also uses abstraction in micro-management like the previous two bots. But it also uses it in macro-management: as can be seen, the production manager sends commands to the building manager, who is in charge of producing the buildings. This bot also uses divide-and-conquer, and tasks like scouting and resource gathering are managed by separate, independent modules.
\item Skynet: makes extensive use of both abstraction and divide-and-conquer. Considering the decision making component of Skynet, we can see a high level module that issues commands to a series of tactics modules. The collection of tactic modules queue {\em tasks} (that are analogous to the abstract actions used in SPAR). Each different task has a specific low level module that knows how to execute it. Thus, Skynet uses a 3 layered abstraction hierarchy, and uses divide-and-conquer in all levels except the highest.
\item SPAR: only uses abstraction. Its high-level module determines the strategy to use, and the tactical decision module divides it into a collection of {\em abstract actions}, that are executed by the lower-level modules.
\item AIUR:  is mainly divide-and-conquer oriented, for  macro as well
  as  for  micro,  with  a  slight  abstraction  on  macro  due  to  a
  SpendManager deciding how to spend and share resources among Base, Production and Construction Managers. At the beginning of a game, the MoodManager initializes a ``mood'' which will influence both micro and macro. For example, micro-management is divided into three independent managers: the {\em Defense Manager}, controlling military units when there is nothing special, the {\em Under Attack Manager}, activated when the opponent is attacking our bases, and the {\em Army Manager}, taking control of units when it is time to attack, following a timing given by the current mood.
\item BTHAI: uses a two-tier abstraction hierarchy, where a collection of high-level modules command a collection of lower-level agents in charge of each of the units. At the high-level, BTHAI uses divide-and-conquer, having multiple high-level modules issuing commands to the lower-level units.
\end{itemize}

@@

One interesting aspect of the seven bots described above is that, while all of them are reactive at the lower level (reactive control), most if not all of them, are scripted at the highest level of abstraction. BTHAI reads build and squad formations from a predefined script, Nova's {\em Strategy Manager} is a predefined finite-state machine, BBQ's construction manager reads the build order from a predefined script, and Skynet's {\em BuildOrder Manager} is basically a predefined script. Such scripts describe the strategy that the bots will use, however, such strategy is always fixed. One could see this pre-scripting as if each bot defined a ``high-level programming language'' to describe Starcraft strategies, and the bots themselves are just interpreters of such strategy. Compared to current approaches for Chess or Go, this scripting seems a rigid and inflexible, but responds to the much higher complexity of the Starcraft game. An interesting exception to that is UAlbertaBot, which uses a search algorithm in the {\em Production Manager} to find near-optimal build orders. Another interesting case is AIUR, that uses a {\em Mood Manager} to randomly pick a mood among five (rush, aggressive, defensive, macro, fast expand), which will influence the build order, strategy and tactics, all of them scripted so far. This mood can change during the game, following the opponent behavior.

@@

In the end, Skynet\footnote{\url{http://code.google.com/p/skynetbot/}} won the competition with its solid Protoss play (results are summarized in Table \ref{tab:aiide2011}). The bot executed one of a small set of
strategies randomly at the start of the match based on the map and the race of the opponent. Skynet would
then amass a medium to large sized army and expand before moving out to attack. Good use of Dragoon range
and kiting micromanagement allowed it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

As the CIG competition games were executed manually due to
a lack of available software (the AIIDE program was not yet
available at that time), the organizers separated the ten entries into
two brackets. In each bracket of 5 bots, a round-robin
tournament was held with 10 repetitions per pairing, resulting
in 40 games per bot.
The 5 maps chosen for the first round were selected from the pool
of well-known league play maps found on the internet:
\emph{(2)MatchPoint 1.3}, \emph{(4)Fighting Spirit 1.3}, 
\emph{iCCupdestination 1.1}, \emph{iCCup gaia}, and 
\emph{iCCup great barrier reef}. Each bot pairing played
on every map twice, with switched starting positions.
